Comparative Evaluation of Spark and Flink Stream Processing

### 2017-02-22 paper/paper.tex
fix name convention of trajectory to position message

\subsection{Data Stream Setup}
Our streaming workloads read input data stream from Kafka. In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircrafts position over time, comprise 22 fields of data such as aircraft ID, date message generated, longitude, latitude, and altitude.
\par In our experiment a data stream producer component reads the ADS-B datasets, then publish it to Kafka cluster to be consumed by the workloads in Spark and Flink. Figure 1 illustrates the data producer and the Kafka cluster setup.
The data stream is portioned into four portions over two server, while the data producer publish the stream records randomly to Kafka partitions.

@@

\par The used Spark Streaming API operations to execute those steps are shown in Figure 2. First, in Spark to consume parallel Kafka stream, multiple Kafka stream receivers must be created then combine them using the stream \texttt{union} API. Second, we parse each stream record (i.e., ADS-B message) then map it to tuple of aircraft ID and aircraft position object through \texttt{mapToPair} transformation API. Third, the irrelevant messages are filtered by the \texttt{filter}. Fourth, the tuples with same ID are grouped together using the \texttt{groupByKey} Spark API function, after applying this transformation the input stream records are shuffled which means to preserve the order of records a sort operation is needed. Finally, due to the mini-batch nature of Spark streaming, the input stream is processed batch by batch, we use the  \texttt{updateStateByKey} to compute statistics per trajectory after sorting the list of positions, then update the aggregated state of the trajectories stream.

@@

The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the trajectories stream. Given that the dataset of sectors (i.e., polygons) available as reference to check against it, also it used to assign the corresponding sector for certain aircraft position. In this workload, providing the  global model of data (sectors dataset) in streaming processing workload is tested.Moreover, preserving the stream records order and parallel streams consuming are covered.

@@

 The Flink implementation of sector change detection workload, uses the \texttt{FlinkKafkaConsumer09} that read the ADS-B messages stream from Kafka, which used to construct \texttt{DataStream} instance. The \texttt{DataStream} records are parsed to tuples of aircraft ID and aircraft position, by a configured \texttt{map} function. The irrelevant tuples are filtered based on the message type using a \texttt{filter} operation. Afterward, the tuples with common IDs are grouped together by the \texttt{keyBy} operator. Since the Flink streaming processes the stream records as it come, a defined \texttt{reduce} operator assign corresponding sector of new tuple, and the old tuple with same ID is used to retrieve the previous sector to attach it to the new tuple, while the old tuple is discarded. Finally, the tuples with different previous and current sector that represent change in sector case, they are filtered using defined \texttt{filter} transformation.
 \end{itemize}
\section{Performance Results}

>>>>

\subsection{Data Stream Setup}
Our streaming workloads read input data stream from Kafka. In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircraft's position over time, comprise 22 fields of data such as aircraft ID, date message generated, longitude, latitude, and altitude.
\par In our experiment a data stream producer component reads the ADS-B datasets, then publish it to Kafka cluster to be consumed by the workloads in Spark and Flink. Figure 1 illustrates the data producer and the Kafka cluster setup.
The data stream is portioned into four portions over two server, while the data producer publish the stream records randomly to Kafka partitions.

@@

\par The used Spark Streaming API operations to execute those steps are shown in Figure 2. First, in Spark to consume parallel Kafka stream, multiple Kafka stream receivers must be created then combine them using the stream \texttt{union} API. Second, we parse each stream record (i.e., ADS-B message) then map it to tuple of aircraft ID and a position of aircraft object through \texttt{mapToPair} transformation API. Third, the irrelevant messages are filtered by the \texttt{filter}. Fourth, the tuples with same ID are grouped together using the \texttt{groupByKey} Spark API function, after applying this transformation the input stream records are shuffled which means to preserve the order of records a sort operation is needed. Finally, due to the mini-batch nature of Spark streaming, the input stream is processed batch by batch, we use the  \texttt{updateStateByKey} to compute statistics per trajectory after sorting the list of positions, then update the aggregated state of the trajectories stream.

@@

The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the trajectories stream. Given that the dataset of sectors (i.e., polygons) available as reference to check against it, also it used to assign the corresponding sector for certain aircraft's position. In this workload, providing the  global model of data (sectors dataset) in streaming processing workload is tested.Moreover, preserving the stream records order and parallel streams consuming are covered.

@@

 The Flink implementation of sector change detection workload, uses the \texttt{FlinkKafkaConsumer09} that read the ADS-B messages stream from Kafka, which used to construct \texttt{DataStream} instance. The \texttt{DataStream} records are parsed to tuples of aircraft ID and aircraft's position, by a configured \texttt{map} function. The irrelevant tuples are filtered based on the message type using a \texttt{filter} operation. Afterward, the tuples with common IDs are grouped together by the \texttt{keyBy} operator. Since the Flink streaming processes the stream records as it come, a defined \texttt{reduce} operator assign corresponding sector of new tuple, and the old tuple with same ID is used to retrieve the previous sector to attach it to the new tuple, while the old tuple is discarded. Finally, the tuples with different previous and current sector that represent change in sector case, they are filtered using defined \texttt{filter} transformation.
 \end{itemize}
\section{Performance Results}
